Price of Anarchy for Mean Field Games

### 2018-08-01 PoA.tex
add theorem 2 equivalent conditions for PoA =1

It will also be useful to compute the derivative of $u_t$ with respect to time $t$ from the explicit form in equation (\ref{eq:riccati_ut_sol}):
\begin{equation}
\frac{d u_t}{dt} = \frac{\left(B (D^{u})^2 + 2 A^u D^{u} - C^{u} \right) \cdot \left(\delta_u^{+} - \delta_u^{-}\right)^2 e^{- (\delta_u^{+} - \delta_u^{-})(T-t) } }{\left[B D^{u}(1-e^{- (\delta_u^{+} - \delta_u^{-})(T-t)})   + \delta_u^{+}e^{- (\delta_u^{+} - \delta_u^{-})(T-t)}- \delta_u^{-}  \right]^2 }.
\label{eq:riccati_ut_sol_deriv}
\end{equation}
Note that $u_t$ is increasing if $B (D^{u})^2 + 2 A^u D^{u} - C^{u}>0$, and likewise, decreasing if $B (D^{u})^2 + 2 A^u D^{u} - C^{u}<0$.\\

@@

	Step 1: We derive a lower bound for $\Delta SC^{b_2}$ by adapting the techniques used in inequality (\ref{eq:I1_ineq_b2_infinity}).  We have shown that every $t \in[0,T]$,  $\lim_{b_2 \to \infty} h_u(b_2, t)= c_u$ and $\lim_{b_2 \to \infty} h_w(b_2, t)= c_w$. Let $c_3 = \ln(2) \frac{r + \bar{r}(1-\bar{s})^2}{T (c_u + \vert c_u - c_w \vert / 4)} - \bar{b}_2$. Then, there exists $b_2^{*,numerator} \geq \max \{ b_2^{*,basic}, c_3 \}$ such that for all $b_2 \geq b_2^{*,numerator}$ and all $s \in [0, T/2]$, we have:
	\begin{equation*}
	\begin{array}{lclclclcl}
		c_u - \left\vert c_u - c_w  \right\vert /4 &\leq& h_u(b_2, 0) &\leq& h_u(b_2,s) &\leq& h_u(b_2, T/2) &\leq& c_u + \left\vert c_u - c_w \right\vert /4, \\
		c_w -  \left\vert c_u - c_w  \right\vert /4 &\leq& h_w(b_2, 0) &\leq& h_w(b_2,s) &\leq& h_w(b_2, T/2) &\leq& c_w + \left\vert c_u - c_w  \right\vert /4,
	\end{array}
	\end{equation*}
	which implies that for all $t \in [0,T/2]$:
	$$ \left\vert h_u(b_2, t) - h_w(b_2,t)\right\vert \geq \frac{1}{2} \left\vert c_u - c_w \right\vert, \qquad \int_0^t h_u(b_2, s) ds \leq \left( c_u + \frac{1}{4} \left\vert c_u - c_w \right\vert \right) t.
	$$
	Thus, similar to inequality (\ref{eq:I1_ineq_b2_infinity}), for all $b_2 \geq b_2^{*,numerator}$, we deduce:
	\begin{IEEEeqnarray}{rCl}
		(b_2 + \bar{b}_2) \Delta SC ^{b_2} &\geq & \frac{b_2 + \bar{b}_2}{2} B^{b_2} \int_0^{\frac{T}{2}} (u^{b_2}_t - w^{b_2}_t)^2 (\bar{x}^{MFG, b_2})^2 dt \nonumber \\
		&\geq& \frac{\left\vert c_u - c_w \right\vert^2 (\mathbb{E}(\xi))^2}{8 (r+ \bar{r}(1-\bar{s})^2 )} \cdot (b_2 + \bar{b}_2) \int_0^{\frac{T}{2}} \exp \left[- \frac{2 (b_2 + \bar{b}_2)}{r +\bar{r}(1-\bar{s})^2} \left( c_u + \frac{1}{4} \left\vert c_u - c_w \right\vert \right) t \right] dt \nonumber \\
		&\geq & \frac{\left\vert c_u - c_w \right\vert^2 (\mathbb{E}(\xi))^2}{16 (c_u +  \vert c_u - c_w \vert / 4)} \left( 1- e^{- \frac{T (c_u + \vert c_u - c_w \vert/4 )}{r+\bar{r}(1-\bar{s})^2}(b_2 + \bar{b}_2)} \right) \nonumber \\
		&\geq & (2 c_{num} ) \cdot \frac{1}{2} = c_{num},
	\label{eq:numerator_ineq_b2_infinity}	
	\end{IEEEeqnarray}
	with $c_{num} =  \frac{\left\vert c_u - c_w \right\vert^2 (\mathbb{E}(\xi))^2}{32 (c_u +  \vert c_u - c_w \vert / 4)}$ and the last inequality is justified because $b_2 \geq c_3$. \\

>>>>

It will also be useful to compute the derivative of $u_t$ with respect to time $t$ from the explicit form in equation (\ref{eq:riccati_ut_sol}):
\begin{equation}
\frac{d u_t}{dt} = \frac{\left(B (D^{u})^2 + 2 A^u D^{u} - C^{u} \right) \cdot \left(\delta_u^{+} - \delta_u^{-}\right)^2 e^{- (\delta_u^{+} - \delta_u^{-})(T-t) } }{\left[B D^{u}(1-e^{- (\delta_u^{+} - \delta_u^{-})(T-t)})   + \delta_u^{+}e^{- (\delta_u^{+} - \delta_u^{-})(T-t)}- \delta_u^{-}  \right]^2 }.
\label{eq:riccati_ut_sol_deriv}
\end{equation}
Note that $u_t$ is increasing if $B (D^{u})^2 + 2 A^u D^{u} - C^{u}>0$, and likewise, decreasing if $B (D^{u})^2 + 2 A^u D^{u} - C^{u}<0$.

\begin{theorem}
	Assuming (\ref{eq:assumptions}), if the initial condition $\xi$ satisfies that $\mathbb{E}(\xi)  \neq 0$, then 
	\begin{itemize}
		\item when $\bar{b}_1 > 0$, we have $PoA = 1$ if and only if
		\begin{equation}
		D^u = D^w =:D \quad \text{and} \quad B D^2 + 2 A^{u} D - C^{u} = B D^2 + 2 A^{w} D - C^w  = 0,
		\label{eq:PoA=1_condition_b1_bar>0}
		\end{equation}
		\item when $\bar{b}_1 = 0$, then $A^u = A^w$ and we have $PoA = 1$ if and only if 
		\begin{equation}
			D^u = D^w \quad \text{and} \quad C^u = C^w, 
		\label{eq:PoA=1_condition_b1_bar=0}
		\end{equation}
	\end{itemize}
	Namely, $PoA = 1$ if and only if the following two or three conditions hold true:
	\begin{equation*}
	\left\{
	\begin{array}{rcl}
	\displaystyle \lambda \cdot (q_T + \bar{q}_T(1-s_T) )  &=& q_T + \bar{q}_T (1- s_T)^2\\
	\displaystyle \lambda \cdot (q+ \bar{q}(1-s)) &=& (q + \bar{q}(1-s)^2) + \bar{b}_1 \cdot (q_T + \bar{q}_T (1 - s_T)^2) 
	\end{array}
	\right.
	\end{equation*}
	where $\lambda = \frac{b_2}{b_2 + \bar{b}_2}\cdot \frac{ r + \bar{r}(1-\bar{s})^2 }{r + \bar{r}(1-\bar{s})}$, and in the case when $\bar{b}_1 > 0$, the third condition becomes
	$$ \frac{(b_2+ \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} ( q_T + \bar{q}_T(1-s_T)^2 )^2 - 2 (b_1 + \bar{b}_1) ( q_T + \bar{q}_T(1-s_T)^2 )  + ( q + \bar{q}(1-s)^2 ) = 0. $$

	\label{thm:PoA_equivalent_1}
\end{theorem}

\begin{proof}
	Under assumption (\ref{eq:assumptions}), we know from equation (\ref{eq:riccati_wt}) and from the explicit formula (\ref{eq:riccati_ut_sol}) for solution to scalar Riccati equation that $w_0$ is strictly positive. Thus, with the assumption $\mathbb{E}(\xi) \neq 0$, we have $SC^{MKV} > 0$. Hence, $PoA =1$ if and only if $\Delta SC = 0$. Since $\bar{x}_t^{MFG} \neq 0$ for all $t \in [0,T]$, from proposition \ref{prop:diff_SC} and the continuity of $u_t$ and $w_t$, we derive that 
	\begin{equation}
		PoA = 1 \qquad iff \qquad u_t = w_t, \quad \forall  t \in [0,T].
	\label{eq:PoA_equivalence_ut=w_t} 
	\end{equation}

	From equation (\ref{eq:riccati_ut_sol_deriv}) and the Riccati equations (\ref{eq:riccati_ut}) and (\ref{eq:riccati_wt}), it is easy to check that if the conditions in (\ref{eq:PoA=1_condition_b1_bar>0}) and (\ref{eq:PoA=1_condition_b1_bar=0}) are satisfied, we must have $u_t = w_t$ for all $ t\in [0,T]$. Thus, $PoA = 1$.

	Suppose now that $PoA =1$. Then by taking $t=T$ in equation (\ref{eq:PoA_equivalence_ut=w_t}), we obtain the first condition in both (\ref{eq:PoA=1_condition_b1_bar>0}) and (\ref{eq:PoA=1_condition_b1_bar=0}):
	$$D^u = u_T = w_T = D^w.$$
	Moreover, if we make the difference between the two Riccati equations (\ref{eq:riccati_ut}) and (\ref{eq:riccati_wt}), and by taking $u_t = w_t$ for all $t in [0,T]$ into account, we obtain:
	\begin{equation}
		2 (A^u - A^w) w_t = C^u - C^w, \quad \forall \  t \in [0,T]
	\label{eq:wt=constant}
	\end{equation}
	Since $2(A^u - A^w) = \bar{b}_1$, so that in the case when $\bar{b}_1 =0$ we must have $C^u = C^w$. Otherwise, equation (\ref{eq:wt=constant}) implies that $u_t = w_t = (C^u - C^w) / \bar{b}_1 $, which are constants for all $t \in [0,T]$. Thus, the derivatives with respect to time $t$ for $u_t$ and $w_t$ should be zero. From equation (\ref{eq:riccati_ut_sol_deriv}), and the fact that $\delta_u^+ - \delta_u^- > 0$, we deduce: 
	$$ B (D^{u})^2 + 2 A^u D^u - C^u = 0.$$
	Similarly we also have $B (D^{w})^2 + 2 A^w D^w - C^w = 0$.

	By replacing $(A^u,B, C^u, D^u)$ and $(A^w, B, C^w, D^w)$ with their explicit expressions, we obtain henceforth an equivalent condition for $PoA = 1$ in term of the coefficients of Linear Quadratic model.

@@

	Step 1: We derive a lower bound for $\Delta SC^{b_2}$ by adapting the techniques used in inequality (\ref{eq:I1_ineq_b2_infinity}).  It has been shown that for every $0 \leq t <T$,  $h_u(b_2, t) \xrightarrow[b_2 \to \infty]{} c_u$ and $h_w(b_2, t) \xrightarrow[b_2 \to \infty]{} c_w$. Let $c_3 = \ln(2) \frac{r + \bar{r}(1-\bar{s})^2}{T (c_u + \vert c_u - c_w \vert / 4)} - \bar{b}_2$. Then, there exists $b_2^{*,numerator} \geq \max \{ b_2^{*,basic}, c_3 \}$ such that for all $b_2 \geq b_2^{*,numerator}$ and all $s \in [0, T/2]$ we have:
	\begin{equation*}
	\begin{array}{lclclclcl}
		c_u - \left\vert c_u - c_w  \right\vert /4 &\leq& h_u(b_2, 0) &\leq& h_u(b_2,s) &\leq& h_u(b_2, T/2) &\leq& c_u + \left\vert c_u - c_w \right\vert /4, \\
		c_w -  \left\vert c_u - c_w  \right\vert /4 &\leq& h_w(b_2, 0) &\leq& h_w(b_2,s) &\leq& h_w(b_2, T/2) &\leq& c_w + \left\vert c_u - c_w  \right\vert /4,
	\end{array}
	\end{equation*}
	which implies that for all $t \in [0,T/2]$
	$$ \left\vert h_u(b_2, t) - h_w(b_2,t)\right\vert \geq \frac{1}{2} \left\vert c_u - c_w \right\vert, \qquad \int_0^t h_u(b_2, s) ds \leq \left( c_u + \frac{1}{4} \left\vert c_u - c_w \right\vert \right) \cdot t.
	$$
	Thus, similar to inequality (\ref{eq:I1_ineq_b2_infinity}), for all $b_2 \geq b_2^{*,numerator}$ we deduce
	\begin{IEEEeqnarray}{rCl}
		(b_2 + \bar{b}_2) \Delta SC ^{b_2} &\geq & \frac{b_2 + \bar{b}_2}{2} B \int_0^{\frac{T}{2}} (u_t - w_t)^2 (\bar{x}^{MFG, b_2})^2 dt \nonumber \\
		&\geq& \frac{\left\vert c_u - c_w \right\vert^2 (\mathbb{E}(\xi))^2}{8 (r+ \bar{r}(1-\bar{s})^2 )} \cdot (b_2 + \bar{b}_2) \int_0^{\frac{T}{2}} \exp \left( - \frac{2 (b_2 + \bar{b}_2)}{r +\bar{r}(1-\bar{s})^2} \cdot \left( c_u + \frac{1}{4} \left\vert c_u - c_w \right\vert \right) \cdot t \right) dt \nonumber \\
		&\geq & \frac{\left\vert c_u - c_w \right\vert^2 (\mathbb{E}(\xi))^2}{16 (c_u +  \vert c_u - c_w \vert / 4)} \left( 1- e^{- \frac{T (c_u + \frac{1}{4} \vert c_u - c_w \vert )}{r+\bar{r}(1-\bar{s})^2}\cdot (b_2 + \bar{b}_2)} \right) \nonumber \\
		&\geq & (2 c_{num} ) \cdot \frac{1}{2} = c_{num}
	\label{eq:numerator_ineq_b2_infinity}	
	\end{IEEEeqnarray}
	where $c_{num} =  \frac{\left\vert c_u - c_w \right\vert^2 (\mathbb{E}(\xi))^2}{32 (c_u +  \vert c_u - c_w \vert / 4)}$ and the last inequality is justified by the choice of $b_2 \geq c_3$. \\
